---
title: "Analysis of the Activity Dataset"
output: html_document
date: "2023-04-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the necessary libraries

```{r, results='hide', message=FALSE}
library(dplyr)
library(ggplot2)
```

## Reading Data 

```{r}
path <- "./data/activity.csv"
df <- read.csv(path)
df$date <- as.Date(df$date)

summary(df)
```


## Histogram of the total number of steps taken each day

```{r}
steps_per_day <- df %>% group_by(date) %>% summarise(steps=sum(steps, na.rm = TRUE))
ggplot(steps_per_day, aes(x=steps)) + geom_histogram(bins = 50)

```

## Mean and median number of steps taken each day

```{r}
df %>%
  group_by(date) %>% 
  summarise(
    steps_mean=mean(steps, na.rm = TRUE), 
    steps_median=median(steps, na.rm = TRUE)
  ) 
```

Since most of intervals in a given day have 0 steps, the median is pretty much
always equal to 0. 


## Time series plot of the average number of steps taken

```{r}
mean_steps <- df %>%
  group_by(date) %>% 
  summarise(steps=mean(steps, na.rm = TRUE)) 

ggplot(mean_steps, aes(x=date, y=steps)) + geom_line()
```

Since we haven't removed the missing data, the timeseries has some gaps.

## The 5-minute interval that, on average, contains the maximum number of steps

```{r}
df %>%
  group_by(interval) %>%
  summarise(mean_steps = mean(steps, na.rm = TRUE)) %>%
  arrange(desc(mean_steps)) %>%
  head(n = 1)
```

The interval 835-840 of minutes on average has the greatest number of steps.


## Code to describe and show a strategy for imputing missing data

```{r}
df[is.na(df$steps),] %>% group_by(date) %>% summarise(n())
```

Looking at the missing data it seems that we're always missing the whole days.

Usually one day at a time. There's one case 09-11-2012 - 10-11-2012 where there 
are two missing days in a row. 

Fill the missing data using the mean value for each  of the intervals:

```{r}
filled_df <- df %>%
  group_by(interval) %>% 
  mutate(steps=ifelse(is.na(steps), mean(steps, na.rm = TRUE), steps))
```

## Histogram of the total number of steps taken each day after missing values are imputed

```{r}
mean_steps <- filled_df %>%
  group_by(date) %>% 
  summarise(steps=mean(steps, na.rm = TRUE)) 

ggplot(mean_steps, aes(x=date, y=steps)) + geom_line()
```

We can see that the new timeseries doesn't have the gaps that the previous
one had.


## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
data <- filled_df %>%
  mutate(is_weekend=format(date, "%a") %in% c("Sat", "Sun")) %>%
  group_by(is_weekend, interval) %>%
  summarise(steps=mean(steps, na.rm = TRUE), .groups = "drop") 

ggplot(data, aes(x=interval, y=steps)) + 
  geom_line() + 
  facet_grid(
    . ~ is_weekend, 
    labeller = labeller(is_weekend = c("TRUE" = "Weekend", "FALSE" = "Weekday"))) +
  labs(title = "Average number of steps across weekdays and weekends")
```

